Comparison of FIA and equally-sized communities

### 2021-05-09 analysis/reports/submission2/appendices/appendix1_filtering.Rmd

---
title: "Filtering datasets"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    word_document:
       df_print: kable
---

@@

1) When communities have only 1 mathematically possible SAD (S = 1, N = S, or N = S + 1).
2) When the sampled feasible sets yield fewer than 20 unique values for skewness/evenness. If there are fewer than 20 unique values in the comparison vector, it is impossible to be in the 5th or 95th percentile.
3) When a community consists of only 2 species, we excluded it from analyses using skewness, because e1071::skewness() always evaluates to 0 if S = 2.

@@

all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

@@

Finally, we filter for cases 2 and 3. That is, for aggregate analyses, we will restrict our analyses to sites whose feasible sets have more than 20 unique values for skewness or evenness (whichever metric we are focused on). For analyses with skewness, we will also remove sites with only 2 species. This results in these final total numbers of communities included for skewness and evenness.

@@

all_di_filtered %>%
  filter(skew_unique > 20, s0 > 2) %>%
  group_by(dat) %>%
  summarize(sites_for_skewness = dplyr::n()) %>%
  mutate(total_sites_for_skewness = sum(sites_for_skewness))

@@

For evenness:

@@

all_di_filtered %>%
  filter(simpson_unique > 20) %>%
  group_by(dat) %>%
  summarize(sites_for_evenness = dplyr::n()) %>%
  ungroup() %>%
  mutate(total_sites_for_evenness = sum(sites_for_evenness))

>>>>

---
title: "Filtering datasets"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    word_document:
       df_print: kable
       reference_docx: style_reference.docx

---

@@

1) When communities have only 1 mathematically possible SAD (S = 1, N = S, or N = S + 1).
2) When there are fewer than 20 (for the dissimilarity metric) or 40 (for all other metrics) possible SADs. In these cases, it is not possible to fall in the 95th or 97.5/2.5th percentiles.
3) When a community consists of only 2 species, we excluded it from analyses using skewness, because e1071::skewness() always evaluates to 0 if S = 2.

@@

all_di <- read.csv(here::here("analysis", "reports", "submission2", "all_di.csv"), stringsAsFactors = F)

@@

Finally, we filter for cases 2 and 3. That is, for aggregate analyses, we will restrict our analyses to sites whose feasible sets have more than 20 unique values for the dissimilarity metric, and 40 for all others. For analyses with skewness, we will also remove sites with only 2 species. This results in these final total numbers of communities included for dissimilarity, skewness, and all other metrics:

For dissimilarity:

all_di_filtered %>%
  filter(nparts > 20) %>%
  group_by(dat) %>%
  summarize(sites_for_diss = dplyr::n()) %>%
  mutate(total_sites_for_diss = sum(sites_for_diss))

@@

all_di_filtered %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(dat) %>%
  summarize(sites_for_skewness = dplyr::n()) %>%
  mutate(total_sites_for_skewness = sum(sites_for_skewness))

@@

For all other metrics:

@@

all_di_filtered %>%
  filter(nparts > 40) %>%
  group_by(dat) %>%
  summarize(sites_for_others = dplyr::n()) %>%
  ungroup() %>%
  mutate(total_sites_for_others = sum(sites_for_others))

### 2021-05-09 analysis/reports/submission2/appendices/s8_gentry.Rmd

---
title: "Unusual Gentry communities"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    # github_document:
    #    df_print: kable
    #    toc: true
   word_document:
        df_print: kable
        toc: false
        reference_docx: style_reference.docx
---

@@

The Gentry communities are extreme relative to hte other datasets (in grey), in that they have high species richness and low abundance. Of these, the most extreme communities have extremely low average abundance (e.g. N/S < 3).

@@

ggplot(gentry_di, aes(s0, n0, color = `Low average \nabundance`)) +
  geom_point(data = all_di, inherit.aes = F, aes(s0, n0), color = "grey", alpha = .1) +  geom_point() +
scale_x_log10() + scale_y_log10() + xlab("Species richness (S); note log scale") + ylab("Total abundance (N); note log scale") + ggtitle("Gentry communities", subtitle = "(in color)") + theme(legend.position = "bottom")

@@

These low-average-abundance communities - the red dots in the plot above - account for most of the signal of extreme values in the unusual direction for various metrics - that is, a **low** number of rare species and **low** skewness, and **high** Shannon and Simpson evenness. Gentry communities with average abundance > 3 tend to show the signal in the same direction as most other communities.

@@

>>>>

---
title: "Unusual Gentry communities"
#author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    # github_document:
    #    df_print: kable
    #    toc: true
   word_document:
        df_print: kable
        toc: false
        reference_docx: style_reference.docx
---

@@

The Gentry communities are extreme relative to the other datasets (in grey), in that they have high species richness and low abundance. Of these, the most extreme communities have extremely low average abundance (e.g. N/S < 3).

@@

ggplot(gentry_di, aes(s0, n0, color = `Low average \nabundance`)) +
  geom_point(data = all_di, inherit.aes = F, aes(s0, n0), color = "grey", alpha = .1) +  geom_point() +
scale_x_log10() + scale_y_log10() + xlab("Species richness (S); note log scale") + ylab("Total abundance (N); note log scale") + ggtitle("Gentry communities", subtitle = "(in color)") + theme(legend.position = "bottom") + scale_color_viridis_d(end = .8)

@@

These low-average-abundance communities - the dark purple dots in the plot above - account for most of the signal of extreme values in the unusual direction for various metrics - that is, a **low** number of rare species and **low** skewness, and **high** Shannon and Simpson evenness. Gentry communities with average abundance > 3 tend to show the signal in the same direction as most other communities.

@@

These low-average abundance communities have unusual statistical baselines. For example, samples from these feasible sets have a very high proportion of rare species:

```{r, fig.dim = c(4.5,4)}

all_di <- all_di %>%
  mutate(`Average proportion of rare species \nin the feasible set` = nsingletons_mean / s0)

ggplot(all_di, aes(s0, n0, color = `Average proportion of rare species \nin the feasible set`)) +
  geom_point(alpha = .4) +
scale_x_log10() + 
  scale_y_log10() + 
  scale_color_viridis_c(option = "plasma", end = .8) +
  xlab("Species richness (S); note log scale") + ylab("Total abundance (N); note log scale") + 
  theme(legend.position = "bottom", plot.margin =unit(c(4,4,4,4), units = "mm"))

### 2021-05-09 analysis/reports/submission2/appendices/s9_fia_comparison.Rmd

```{r, fig.dim = c(8,4)}
fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_percentile_hist(sub_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(sub_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(sub_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(sub_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(sub_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))

@@

```{r, fig.dim = c(8,4)}
fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_narrowness_hist(sub_di, "sim_pos_from_best_95", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_narrowness_hist(sub_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
 left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white"))
```

>>>>

```{r, fig.dim = c(8,8)}
fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_percentile_hist(sub_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(sub_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(sub_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(sub_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(sub_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 3,
left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))

@@

```{r, fig.dim = c(8,8)}
fig_2 <- gridExtra::grid.arrange(grobs = list(

  plot_narrowness_hist(sub_di, "sim_pos_from_best_95", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_narrowness_hist(sub_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"),
  plot_narrowness_hist(sub_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset")
), ncol = 3,
left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white"))
```
